---
title: "CI for fWAR"
author: "Rob Weber"
date: "July 30, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Confidence Interval for fWAR
First, we need to find the stretch of years in a player's career that have, on average, the least amount variance. We will then sample randomly from a given number of these years and use combonatorial theory to find a confidence interval

## Data
In this Rmd, "df" will be the full 50 year fangraphs data set and "rook_67" is the data set of the names of rookies in 1967.

## Cleaning
```{r Cleaning}
# get the names from a factor to a character
as.character(df$Name) -> df$Name 

# get the year of their career (kind of, for non-rookies in 67 it's weird right now)
df <- df %>%
  arrange(Season) %>%
  group_by(Name) %>%
  mutate(Season_Num = c(1:length(Name))) %>%
  ungroup()
# to get rid of player's half-careers, need to find all non-rookies in 1967
## all the names from 1967
df$Name[df$Season == 1967] -> name_67
as.character(name_67) -> name_67
## the names of the rookies from 1967
rook_67$Name -> rook_67
as.character(rook_67) -> rook_67
## now we need to get a list of all the player's whose career didn't start in 1967
for(i in 1:length(name_67)) {
  if(name_67[i] %in% rook_67) {
    NA -> name_67[i]
  }
}
## now get rid of the NAs I created
na.omit(name_67) -> name_67
as.character(name_67) -> name_67
## next is to get rid of all of the remaining players from the data set
for(i in 1:length(name_67)) {
  df[which(df$Name != name_67[i]), ] -> df
}
```
## Investigation into the stretch of player's careers with the least variance from year to year.
```{r Variance Investigation}
# now, let's narrow it down to players with at least 3 seasons on record
qual_names <- df %>%
  group_by(Name) %>%
  summarise(Seasons = max(Season_Num)) %>%
  filter(Seasons >= 3)
qual_names$Name -> qual_names
df <- df %>%
  filter(Name %in% qual_names)
# placeholder for the length of the career of the player in question
career_length = 0
# data frame to store results
data.frame("Year_1" = 0, "Year_2" = 0, "Difference" = 0) -> var.df
# data frame to store results for each player inside the loop
for(i in qual_names) {
  # need to grab all the rows of data for the player in question
  df[which(df$Name == i), ] -> df_player
  # need to get the length of his career 
  max(df_player$Season_Num) -> career_length
  # now we need to get a df that we can grab two different years out of
  data.frame("Year_1" = c(1:(career_length - 1)), "Year_2" = c(2:career_length), "Difference" = 0) -> var.df_player
  # now to get the difference in WARs between the two years specified
  for(j in c(1:length(var.df_player$Year_1))) {
    abs(df_player$WAR[df_player$Season_Num == var.df_player$Year_1[j]] - 
          df_player$WAR[df_player$Season_Num == var.df_player$Year_2[j]]) -> var.df_player$Difference[j]
  }
  # now to combine this with the rest of the data
  rbind(var.df, var.df_player) -> var.df
}
# have to get rid of that empty 1st row quick
var.df[which(var.df$Year_1 != 0), ] -> var.df

# now to check which chunks of time have the least variance 
## need
df.sep <- var.df %>%
  group_by(Year_1, Year_2) %>%
  summarise(Avg_Difference = mean(Difference)) 
ggplot(df.sep, aes(x = Year_2, y = Avg_Difference)) +
  geom_point()
## year 5 to year 10 is pretty obviously the most consistent years
```
## Computation
```{r Computation}
# first we need players that played at least 10 years, so let's use the same method as earlier
qual_names <- df %>%
  group_by(Name) %>%
  summarise(Seasons = max(Season_Num)) %>%
  filter(Seasons >= 10)
qual_names$Name -> qual_names
df <- df %>%
  filter(Name %in% qual_names)
# now to just get years 5 through 10
df <- df %>%
  filter(Season_Num >= 5 & Season_Num <= 10)

# next, we need to normalize the WARs in respect to the season number
## this gets the average WAR at each season number 5-10
year_means <- df %>%
  group_by(Season_Num) %>%
  summarise(Mean = mean(WAR))
## this gets the mean that can be used to adjust and center the player WARs
year_means$Mean_adj <- year_means$Mean / mean(df$WAR)
# just gets the respective adjusting constant to each player's row that corresponds to their season number
left_join(df, year_means[ , c(1, 3)], by = "Season_Num") -> df

# This gets a column of player's normalized WAR values in respect to season number
df <- df %>%
  group_by(Season_Num) %>%
  mutate(WAR_normal = WAR / Mean_adj)
# now, to get a confidence interval, we will use a manual sampling method to find a 94% CI through combinatorial probability calculations under the assumption that 3 random years of the 6 possible years are being taken and the average is being found.
df.ci <- df %>%
  arrange(Name, WAR_normal) %>%
  group_by(Name) %>%
  summarise(CI_value = (WAR_normal[5] - WAR_normal[2]) / 2) 

mean(df.ci$CI_value)
```
# The result is that the confidence interval for position player fWAR is +/- 0.87
